---
title: "STAT5147 Module3 - Practice"
author: "Aheer Sravon"
format: pdf
editor: visual
---

# Packages and Functions

```{r}
source("./functions.R")
```

# Module 3 Part 2

## Setting Up Discrete Random Variables

```{r}
y <- c(2:12)
proby <- c(1,2,3,4,5,6,5,4,3,2,1)/36
proby
```

The graph of the distribution of the sum of two dice is,

```{r}
plot.discrete(y, proby)
```

Mean of $Y$,

```{r}
sum(y*proby)
```

We can also calculate expected values of functions of $Y$,

```{r}
sum(log(y)*proby)
sum(sqrt(y)*proby)
```

The variance($E[Y^2] - \mu^2$) can be calculated in the following way,

```{r}
sum((y^2)*proby) - sum(y*proby)^2
```

# Module 3 Part 3

## Calculating Probabilities of Binomial Distribution

- $P(Y = y)$ is `dbiom(y, n, p)`
- $P(Y \leq y)$ is `dbiom(0:y, n, p)`
- $P(Y \geq y)$ is `dbiom(y:n, n, p)`

Suppose the random variable $Y$ is the number of times the sum of two dice equals 7 in 30 rolls.

- $P(Y = 8)$
```{r}
dbinom(8, 30, 1/6)
```
- $P(Y \leq 2)$
```{r}
sum(dbinom(0:2, 30, 1/6))
```
- $P(Y \geq 6)$
```{r}
sum(dbinom(6:30, 30, 1/6))
```

Below is the graph of the PMF of the Binomial Random Variable with $n = 30$ and $p = \frac{1}{6}$ -
```{r}
plot.binom(30, 1/6)
```

 # Module 3 Part 4

## Calculating Probabilities of Poisson Distribution

- $P(Y = y)$ is `dpois(y, lambda)`.
- $P(Y \leq y)$ is `sum(dpois(0:y, lambda))`
- $P(Y \geq y) = 1 - P(Y \leq y-1)$

Let $Y$ be the number of bacteria colonies of a certain type in a sample of polluted water. $Y$ has a poisson distribution with a mean of 2 per 1-cubic centimeter.

- $P(Y = 4)$
```{r}
dpois(4, 2)
```

- $P(Y \leq 2)$
```{r}
sum(dpois(0:2, 2))
```

- $P(Y \geq 6)$
```{r}
1 - sum(dpois(0:6-1, 2))
```

```{r}
plot.pois(2)
```


