---
title: "STAT5147 Module 3 Homework"
author: "Aheer Sravon"
format: pdf
editor: visual
---

# Used Function(s)

```{r}
require(ggplot2)

plot.discrete <- function(y, proby) {
bars <- as.data.frame(cbind(y, proby))

ggplot(bars, aes(x=y, y=proby)) +
  geom_bar(stat="identity", width=1, fill="blue", colour="black") +
  theme_bw() +
  theme(axis.title.y=element_text(size=rel(1.4)),
        axis.title.x=element_text(size=rel(1.4)),
        axis.text.x=element_text(size=rel(1.2)),
        axis.text.y=element_text(size=rel(1.2)),
        plot.title=element_text(hjust=0.5, size=rel(1.6))) +
  labs(
    x="y",
    y="p(y)",
    title="Distribution of Discrete R.V."
  )
}

plot.binom <- function(n, p) {
  y <- c(0:n)
  proby <- dbinom(0:n, n, p)
  bars <- as.data.frame(cbind(y, proby))
  
  ggplot(bars, aes(x=y, y=proby)) +
    geom_bar(stat="identity", width=1, fill="blue", colour="black") +
    theme_bw() +
    theme(axis.title.y=element_text(size=rel(1.4)),
          axis.title.x=element_text(size=rel(1.4)),
          axis.text.x=element_text(size=rel(1.2)),
          axis.text.y=element_text(size=rel(1.2)),
          plot.title=element_text(hjust=0.5, size=rel(1.6))) +
    labs(x="Number of Successes (y)",
         y="Probability",
         title="Distribution of Binomial R.V.")
}

plot.pois<- function(lambda){
  maxy<- qpois(0.9999, lambda)
  y<- c(0:maxy)
  proby<- dpois(0:maxy, lambda)
  Bars<- as.data.frame(cbind(y, proby))
  ggplot(Bars, aes(x = y, y = proby))+
    geom_bar(stat="identity", width = 1,
             fill = "blue", colour = "black")+
    theme_bw()+
    theme(axis.title.y = element_text(size = rel(1.4)),
          axis.title.x = element_text(size = rel(1.4)),
          axis.text.x = element_text(size = rel(1.2)),
          axis.text.y = element_text(size = rel(1.2)),
          plot.title = element_text(hjust=0.5, size = rel(1.6)))+
    labs(x = "Number of Events (y)",
         y = "Probability",
         title = "Distribution of Poisson R.V.")
}
```

# Problem 5

Answer to the question e is calculated below,

```{r}
sum(dbinom(60:100, 100, 1/2))
```

Which is only about 3%

# Problem 7

- $p = 0.5, n = 5$

```{r}
plot.binom(5, 0.5)
```

- $p = 0.5, n = 10$

```{r}
plot.binom(10, 0.5)
```

- $p = 0.5, n = 20$

```{r}
plot.binom(20, 0.5)
```

- $p = 0.5, n = 50$

```{r}
plot.binom(50, 0.5)
```

- $p = 0.1, n = 25$

```{r}
plot.binom(25, 0.1)
```

- $p = 0.1, n = 50$

```{r}
plot.binom(50, 0.1)
```

- $p = 0.1, n = 100$

```{r}
plot.binom(100, 0.1)
```

- $p = 0.1, n = 250$

```{r}
plot.binom(250, 0.1)
```

- $p = 0.9, n = 25$

```{r}
plot.binom(25, 0.9)
```

- $p = 0.9, n = 50$

```{r}
plot.binom(50, 0.9)
```

- $p = 0.9, n = 100$

```{r}
plot.binom(100, 0.9)
```

- $p = 0.9, n = 250$

```{r}
plot.binom(250, 0.9)
```

So,

a)  $p = 0.5$
b)  $p = 0.1$ and $p = 0.9$ are mirror images of each other for the same sample size $n$.
c)  $p = 0.1$
d)  $p = 0.9$
e)  For the last two distributions, the graphs are approximately symmetric for each value of $p$. As the smaller of $np$ and $n(1-p)$ increases, the binomial distribution becomes less skewed and more symmetric about its mean.

# Problem 8

Poisson distribution with $\lambda = 5$ per hour.

a) Probability that exactly four arrivals occur during a particular hour is,
```{r}
dpois(4, 5)
```

b) Probability that at most 2 people arrive during a particular hour is,
```{r}
sum(dpois(0:2, 5))
```

c) Probability that at least 4 people arrive during a particular hour,
```{r}
1 - sum(dpois(0:3, 5))
```

d) Both the mean and the variance of the number of people I expect to arrive during a particular hour is $5$.

e) It's a binomial distribution with success probability equal to `1 - sum(dpois(0:6, 5))`. So the probability is,
```{r}
(1 - dbinom(0, 8, (1-sum(dpois(0:6, 5)))))
```

# Problem 9

- $\lambda=1$
```{r}
plot.pois(1)
```
- $\lambda=5$
```{r}
plot.pois(5)
```

- $\lambda=25$
```{r}
plot.pois(25)
```

- $\lambda=50$
```{r}
plot.pois(50)
```

As $\lambda$ increases, the Poisson distribution shifts to the right, becomes more spread out, and changes from right-skewed to increasingly symmetric around its mean $\lambda$.